On The Subspace Learning for Network Attack Detection

### 2019-10-17 docs/thesis/parts/abstract.tex
Revert "abstract"
This reverts commit 759f3f56d8e8e954dd6810d9ccfe4ee590c8eac6.

Signal processing techniques are usually applied for adaptive, streaming and real time problems, with high data volume and throughput. These techniques are widely adotped for denoising, signal separation and anomaly detection, and has attracted attention to apply signal processing for network attack detection. Recently, signal processing schemes have been applied to detect malicious traffic in computer networks, showing advances in network traffic analysis and network attack detection. Most of network-wide anomaly traffic detection methods are PCA-based method, which draw on PCA to divide traffic into normal and abnormal subspace, in accordance to anomaly detection problems.

@@

As a first important contribution, we propose the Eigensimilarity, which is an approach based on signal processing concepts applied to detection of malicious traffic in computer networks. Eigensimilarity models the network traffic using a signal processing formulation as a composition of three components: signal, artifact and noise, taking into account the incoming and outgoing traffic in certain types of network ports (TCP or UDP). The Eigensimilarity is based on eigenvalue analysis, model order selection (MOS) and similarity analysis between eigenvectors, where MOS and eigenvalue analysis are applied to detect time frames under attack, while the similarity analysis between eigenvectors aims the detection of the exxact time and network ports under attack. We evaluate the accuracy and performance of the proposed framework applied to a simulated scenario and to the DARPA 1998 data set. The performed experiments show that synflood, fraggle and port scan attacks can be detected accurately by Eigensimilarity and with great detail in an automatic and blind fashion, i.e. in an unsupervised approach.

The high availability of raw data increases the challenges related to big data analytics and to imbalanced learning problems, which corresponds to data sets exhibiting significant imbalances of classes or rare events of some classes. The fundamental issue with the imbalanced learning is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms. Data analysis of imbalanced data is challenging for classification and prediction problems related to anomaly detection, novelty detection, fraud detection and attack detection... link to fraud and skewed data

>>>>

Traditionally, cyber defense methods can be effective against well known attacks, yet may fail against innovative malicious techniques. In order to be able to detect and avoid novel attacks and their variations, it is necessary to develop or improve techniques to achieve efficiency in attack detection, behavioral analysis, resource consumption, processing capacity and response time. Moreover, it is crucial to obtain high detection accuracy and capacity to detect malicious patterns without previous knowledge.

Recently, signal processing schemes have been applied to detect malicious traffic in computer networks, showing advances in network traffic analysis and network attack detection. Signal processing techniques are usually applied for adaptive, streaming and real time problems, with high data volume and throughput. These techniques are widely adotped for denoising, signal separation and anomaly detection, and has attracted attention to apply signal processing for network attack detection. Most of network-wide anomaly traffic detection methods are PCA-based method, which draw on PCA to divide traffic into normal and abnormal subspace, in accordance to anomaly detection problems.

@@

As a first important contribution, we propose the Eigensimilarity, which is an approach based on signal processing concepts applied to detection of malicious traffic in computer networks. Eigensimilarity models the network traffic using a signal processing formulation as a composition of three components: signal, artifact and noise, taking into account the incoming and outgoing traffic in certain types of network ports (TCP or UDP). The Eigensimilarity is based on eigenvalue analysis, model order selection (MOS) and similarity analysis between eigenvectors, where MOS and eigenvalue analysis are applied to detect time frames under attack. We evaluate the accuracy and performance of the proposed framework applied to a simulated scenario and to the DARPA 1998 data set. The performed experiments show that synflood, fraggle and port scan attacks can be detected accurately by Eigensimilarity and with great detail in an automatic and blind fashion, i.e. in an unsupervised approach.

Recent developments in science and technology have enabled the growth and availability of raw data to occur at an explosive rate. This has created an immense opportunity for knowledge discovery and data engineering research to play an essential role in a wide range of applications from daily civilian life to national security. Despite the actual high availability of information, the relevant information of some observations is generally of under much reduced dimensionality compared to available data sets. The extraction of relevant information by identifying the generating causes within classes of signals is useful for security analysis and for classification or anomaly detection problems.

The high availability of raw data increases the challenges related to big data analytics and to imbalanced learning problems, which corresponds to data sets exhibiting significant imbalances of classes or rare events of some classes. The fundamental issue with the imbalanced learning is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms. Data analysis of imbalanced data is challenging for classification and prediction problems related to anomaly detection, novelty detection, fraud detection and attack detection.
